### 2010-06-30 papers/icfp-2010-desc/paper_desc.tex
Desc@ICFP: Fix typos.

In dependently typed languages, $\Sigma$-types can be interpreted as
two different generalisations. This duality is reflected in the
notation we can find in the literature. The notation
$\blue{\Sigma}_{\x : \M{A}} (\M{B}\: \x)$ stresses that
$\Sigma$-types are `dependent sums', generalising of sums over
arbitrary arities, where simply typed languages have finite sums.

>>>>

In dependently typed languages, $\Sigma$-types can be interpreted as
two different generalisations. This duality is reflected in the
notation we can find in the literature. The notation
$\blue{\Sigma}_{\x : \M{A}} (\M{B}\: \x)$ stresses that
$\Sigma$-types are `dependent sums', generalising sums over
arbitrary arities, where simply typed languages have finite sums.

### 2010-06-30 papers/icfp-2010-desc/paper_desc_levitation.tex
Desc@ICFP: Fix typos.

The definition \(\Desc \mapsto \Mu{\DescD}\) is circular,
but the offensive recursion is concealed by a prestidigitation.
Expanding be \(\toDesc{-}\) and propagating types as in
Figure~\ref{fig:type-checking} reveals the
awful truth:
\[
\Desc\mapsto
\Mu{\stk{(\DSigma{\EnumT{\sqr{ \DUnit\:\: \SYMBDSigma\:\: \SYMBDIndx\:\: \SYMBDHindx}}\\}
    {\;\SYMBswitch\:\sqr{ \DUnit\:\: \SYMBDSigma\:\: \SYMBDIndx\:\: \SYMBDHindx}\:(\LAM{\_}\Desc)\\
\;\sqr{\stkm{
\DUnit                                            \\
                                   \DSigma{\Set}{\LAM{\V{S}} \DHindx{\V{S}}{\DUnit}}   \\
                                   \DIndx{\DUnit}                                    \\
                                   \DSigma{\Set}{\LAM{\_} \DIndx{\DUnit}}
}}})}}
\]
The recursion shows up only because we must specify the return type
of the general-purpose \(\F{switch}\), and it is computing a \(\Desc\)!
Although type propagation
allows us to hide this detail \emph{when defining a function}, we cannot
readily suppress this information and check types
when \(\F{switch}\) is fully applied.

@@

The catamorphism is defined by induction on the description $\V{D}$,
with a readily propagated non-dependent return type $\V{T}$.
Given a node $\V{xs}$
and the induction hypotheses, the method ought to build an element of
$\V{T}$. Provided that we know how to make an element of
$\descop{\V{D}}{\V{T}}$, this step will be performed by the algebra
$\V{f}$. Let us take a look at this jigsaw:
%
\[\stk{
\cata : \PITEL{\V{D}}{\Desc}
           \PI{\V{T}}{\Set}
           (\descop{\V{D}}{\V{T}} \To \V{T}) \To 
           \Mu{\V{D}} \To \V{T} \\
\cata\: \V{D}\: \V{T}\: \V{f} \mapsto
  \sind \LAM{\V{xs}}\LAM{\V{hs}} \V{f}\: \SHED
}\]
%
We are left with filling the hole. Recall that we have
\(\Bhab{\V{xs}}{\descop{\V{D}}{\Mu{\V{D}}}}\) and
\(\Bhab{\V{hs}}{\All{\V{D}}{(\Mu{\V{D}})}{(\LAM{\_} \V{T})}{\V{xs}}}\)
at hand. Our goal is to make an element of
\(\descop{\V{D}}{\V{T}}\). Intuitively, $\V{xs}$ is of the right
shape, but its sub-elements are of the wrong type. On the other hand,
for each sub-element of $\V{xs}$, $\V{hs}$ gives us the corresponding
element in $\V{T}$.  Therefore, to construct an element of
\(\descop{\V{D}}{\V{T}}\), we must replace the recursive components of
\(\V{xs}\) by their counterparts from \(\V{hs}\). Let us write a
program to do that---please forgive us if we lapse to a pattern matching
notation, for readability's sake.
%
\[\stk{
\F{replace} : \stk{\PITEL{\V{D}}{\Desc}
                   \PITEL{\V{X},\V{Y}}{\Set}\\
                   \PI{\V{xs}}{\descop{\V{D}}{\V{X}}} 
                   \All{\V{D}}{\V{X}}{(\LAM{\_}\V{Y})}{\V{xs}} \To
                   \descop{\V{D}}{\V{Y}}} \\
\F{replace}\: \DUnit\:          \V{X}\: \V{Y}\: \Void\:          \Void          \mapsto 
    \Void                                                                                                         \\
\F{replace}\: (\DSigma{\V{S}}{\V{D}})\: \V{X}\: \V{Y}\: \pair{\V{s}}{\V{d}}{}\: \V{d'}             \mapsto
    \pair{\V{s}}{\F{replace}\: (\V{D}\: \V{s})\: \V{X}\: \V{Y}\: \V{d}\: \V{d'}}{}                                \\
\F{replace}\: (\DIndx{\V{D}})\:     \V{X}\: \V{Y}\: \pair{\V{x}}{\V{d}}{}\: \pair{\V{y}}{\V{d'}}{} \mapsto        \\
\qquad  \pair{\V{y}}{\F{replace}\: \V{D}\: \V{X}\: \V{Y}\: \V{d}\: \V{d'}}{}                                      \\
\F{replace}\: (\DHindx{\V{H}}{\V{D}})\: \V{X}\: \V{Y}\: \pair{\V{f}}{\V{d'}}{}\: \pair{\V{g}}{\V{d'}}{} \mapsto   \\
\qquad  \pair{\V{g}}{\F{replace}\: \V{D}\: \V{X}\: \V{Y}\: \V{d}\: \V{d'}}{}
}\]
%
Filling the hole in $\F{cata}$ with \(\F{replace}\: \V{D}\:
(\Mu{\V{D}})\: \V{T}\: \V{xs}\: \V{hs}\) closes the problem. In the
type theory, we have built a generic catamorphism. Any datatype will
now come equipped with this operation, for free.

@@

In this section, we will turn to a more ambitious generic operation on
datatype. Given a functor, represented as a tagged description, we
build the free monad over this functor.

@@

Of course, we must equip the resulting datatypes with operations
delivering a monadic interface. As expected, \(\LAM{\x}\DVar{\x}\)
plays the r\^ole of \return, embedding variables into terms. The
\bind\ operation corresponds to \emph{substitution}. We will now
implement it, as a generic function.

@@

Our implementation will appeal to the $\cata$ function developed
previously. So, let us write down the types, and fill as much
arguments to $\cata$ as possible:
%
\[\stk{
\begin{array}{@{}ll}
\subst : & \PITEL{\V{D}}{\TagDesc}
           \PI{\V{X}, \V{Y}}{\Set} 
           (\V{X} \To \Mu{(\toDesc{(\FreeMonad{\V{D}}{\V{Y}})})}) \To \\
         & \Mu{(\toDesc{(\FreeMonad{\V{D}}{\V{X}})})} \To
           \Mu{(\toDesc{(\FreeMonad{\V{D}}{\V{Y}})})} 
\end{array} \\
\subst\: \V{D}\: \V{X}\: \V{Y}\: \V{\sigma} \mapsto
  \cata\: (\toDesc{(\FreeMonad{\V{D}}{\V{X}})})\: 
          (\Mu{(\toDesc{(\FreeMonad{\V{D}}{\V{Y}})})})\: 
          \SHED
}\]
%
We are left with implementing the algebra of the
catamorphism. Intuitively, its role is to catch appearances of
$\DVar{\V{x}}$ and replace them by $\V{\sigma}\: \V{x}$. This
corresponds to the following definition:
%
\[\stk{
\begin{array}{@{}ll}
\apply : & \PITEL{\V{D}}{\TagDesc} 
           \PI{\V{X}, \V{Y}}{\Set} 
           (\V{X} \To \Mu{(\toDesc{\FreeMonad{\V{D}}{\V{X}}})}) \To \\
         & \descop{\toDesc{\FreeMonad{\V{D}}{\V{X}}}}{\Mu{(\toDesc{\FreeMonad{\V{D}}{\V{Y}}})}} \To
           \Mu{(\toDesc{\FreeMonad{\V{D}}{\V{Y}}})}
\\
\end{array} \\
\begin{array}{@{}l@{\:\mapsto\:\:}l}
\apply\: \V{D}\: \V{X}\: \V{Y}\: \V{\sigma}\: \pair{\SYMBDVar}{\V{x}}{}   & \V{\sigma}\: \V{x}                   \\
\apply\: \V{D}\: \V{X}\: \V{Y}\: \V{\sigma}\: \pair{\V{c}}{\V{xs}}{} & \Con{\pair{\V{c}}{\V{xs}}{}}
\end{array}
}\]

@@

Filling the sub-goal with $\apply\: \V{D}\: \V{X}\: \V{Y}\:
\V{\sigma}$ completes the implementation. To sum up, we have
implemented the free monad construction for an arbitrary tagged
description. This gives the developer the ability, for any datatype,
to extend it with a notion of variable. Then, we have equipped this
structure with the corresponding monadic operation, \bind\ and
\return. This construction is an example of type-indexed
datatype~\cite{hinze:generic-haskell}, as found in Generic Haskell:
from a datatype, we build a new datatype and equip it with its
structure.

>>>>

The definition \(\Desc \mapsto \Mu{\DescD}\) is circular,
but the offensive recursion is concealed by a prestidigitation.
Expanding \(\toDesc{-}\) and propagating types as in
Figure~\ref{fig:type-checking} reveals the
awful truth:
\[
\Desc\mapsto
\Mu{\stk{(\DSigma{\EnumT{\sqr{ \DUnit\:\: \SYMBDSigma\:\: \SYMBDIndx\:\: \SYMBDHindx}}\\}
    {\;\SYMBswitch\:\sqr{ \DUnit\:\: \SYMBDSigma\:\: \SYMBDIndx\:\: \SYMBDHindx}\:(\LAM{\_}\Desc)\\
\;\sqr{\stkm{
\DUnit                                            \\
                                   \DSigma{\Set}{\LAM{\V{S}} \DHindx{\V{S}}{\DUnit}}   \\
                                   \DIndx{\DUnit}                                    \\
                                   \DSigma{\Set}{\LAM{\_} \DIndx{\DUnit}}
}}})}}
\]
The recursion shows up only because we must specify the return type
of the general-purpose \(\F{switch}\), and it is computing a \(\Desc\)!
Although type propagation
allows us to hide this detail \emph{when defining a function}, we cannot
readily suppress this information and check types
when \(\F{switch}\) is fully applied.

@@

The catamorphism is defined by induction on the description $\V{D}$,
with a readily propagated non-dependent return type $\V{T}$.
Given a node $\V{xs}$
and the induction hypotheses, the method ought to build an element of
$\V{T}$. Provided that we know how to make an element of
$\descop{\V{D}}{\V{T}}$, this step will be performed by the algebra
$\V{f}$. Let us take a look at this jigsaw:
%
\[\stk{
\cata : \PITEL{\V{D}}{\Desc}
           \PI{\V{T}}{\Set}
           (\descop{\V{D}}{\V{T}} \To \V{T}) \To 
           \Mu{\V{D}} \To \V{T} \\
\cata\: \V{D}\: \V{T}\: \V{f} \mapsto
  \sind \LAM{\V{xs}}\LAM{\V{hs}} \V{f}\: \SHED
}\]
%
We are left with filling the hole. Recall that we have
\(\Bhab{\V{xs}}{\descop{\V{D}}{\Mu{\V{D}}}}\) and
\(\Bhab{\V{hs}}{\All{\V{D}}{(\Mu{\V{D}})}{(\LAM{\_} \V{T})}{\V{xs}}}\)
at hand. Our goal is to make an element of
\(\descop{\V{D}}{\V{T}}\). Intuitively, $\V{xs}$ is of the right
shape, but its sub-elements are of the wrong type. On the other hand,
for each sub-element of $\V{xs}$, $\V{hs}$ gives us the corresponding
element in $\V{T}$.  Therefore, to construct an element of
\(\descop{\V{D}}{\V{T}}\), we must replace the recursive components of
\(\V{xs}\) by their counterparts from \(\V{hs}\). Let us write a
program to do that---please forgive us if we lapse to a pattern matching
notation, for readability's sake.
%
\[\stk{
\F{replace} : \stk{\PITEL{\V{D}}{\Desc}
                   \PITEL{\V{X},\V{Y}}{\Set}\\
                   \PI{\V{xs}}{\descop{\V{D}}{\V{X}}} 
                   \All{\V{D}}{\V{X}}{(\LAM{\_}\V{Y})}{\V{xs}} \To
                   \descop{\V{D}}{\V{Y}}} \\
\F{replace}\: \DUnit\:          \V{X}\: \V{Y}\: \Void\:          \Void          \mapsto 
    \Void                                                                                                         \\
\F{replace}\: (\DSigma{\V{S}}{\V{D}})\: \V{X}\: \V{Y}\: \pair{\V{s}}{\V{d}}{}\: \V{d'}             \mapsto
    \pair{\V{s}}{\F{replace}\: (\V{D}\: \V{s})\: \V{X}\: \V{Y}\: \V{d}\: \V{d'}}{}                                \\
\F{replace}\: (\DIndx{\V{D}})\:     \V{X}\: \V{Y}\: \pair{\V{x}}{\V{d}}{}\: \pair{\V{y}}{\V{d'}}{} \mapsto        \\
\qquad  \pair{\V{y}}{\F{replace}\: \V{D}\: \V{X}\: \V{Y}\: \V{d}\: \V{d'}}{}                                      \\
\F{replace}\: (\DHindx{\V{H}}{\V{D}})\: \V{X}\: \V{Y}\: \pair{\V{f}}{\V{d}}{}\: \pair{\V{g}}{\V{d'}}{} \mapsto   \\
\qquad  \pair{\V{g}}{\F{replace}\: \V{D}\: \V{X}\: \V{Y}\: \V{d}\: \V{d'}}{}
}\]
%
Filling the hole in $\F{cata}$ with \(\F{replace}\: \V{D}\:
(\Mu{\V{D}})\: \V{T}\: \V{xs}\: \V{hs}\) closes the problem. In the
type theory, we have built a generic catamorphism. Any datatype will
now come equipped with this operation, for free.

@@

In this section, we will turn to a more ambitious generic operation on
datatypes. Given a functor, represented as a tagged description, we
build the free monad over this functor.

@@

Of course, we must equip the resulting datatypes with operations
delivering a monadic interface. As expected, \(\LAM{\x}\DVar{\x}\)
plays the r\^ole of \return, embedding variables into terms. The
\bind operation corresponds to \emph{substitution}. We will now
implement it, as a generic function.

@@

Our implementation will appeal to the $\cata$ function developed
previously. So, let us write down the types, and fill as many
arguments to $\cata$ as possible:
%
\[\stk{
\begin{array}{@{}ll}
\subst : & \PITEL{\V{D}}{\TagDesc}
           \PI{\V{X}, \V{Y}}{\Set} 
           (\V{X} \To \Mu{(\toDesc{(\FreeMonad{\V{D}}{\V{Y}})})}) \To \\
         & \Mu{(\toDesc{(\FreeMonad{\V{D}}{\V{X}})})} \To
           \Mu{(\toDesc{(\FreeMonad{\V{D}}{\V{Y}})})} 
\end{array} \\
\subst\: \V{D}\: \V{X}\: \V{Y}\: \V{\sigma} \mapsto
  \cata\: (\toDesc{(\FreeMonad{\V{D}}{\V{X}})})\: 
          (\Mu{(\toDesc{(\FreeMonad{\V{D}}{\V{Y}})})})\: 
          \SHED
}\]
%
We are left with implementing the algebra of the
catamorphism. Intuitively, its role is to catch appearances of
$\DVar{\V{x}}$ and replace them by $\V{\sigma}\: \V{x}$. This
corresponds to the following definition:
%
\[\stk{
\begin{array}{@{}ll}
\apply : & \PITEL{\V{D}}{\TagDesc} 
           \PI{\V{X}, \V{Y}}{\Set} 
           (\V{X} \To \Mu{(\toDesc{\FreeMonad{\V{D}}{\V{Y}}})}) \To \\
         & \descop{\toDesc{\FreeMonad{\V{D}}{\V{X}}}}{\Mu{(\toDesc{\FreeMonad{\V{D}}{\V{Y}}})}} \To
           \Mu{(\toDesc{\FreeMonad{\V{D}}{\V{Y}}})}
\\
\end{array} \\
\begin{array}{@{}l@{\:\mapsto\:\:}l}
\apply\: \V{D}\: \V{X}\: \V{Y}\: \V{\sigma}\: \pair{\SYMBDVar}{\V{x}}{}   & \V{\sigma}\: \V{x}                   \\
\apply\: \V{D}\: \V{X}\: \V{Y}\: \V{\sigma}\: \pair{\V{c}}{\V{xs}}{} & \Con{\pair{\V{c}}{\V{xs}}{}}
\end{array}
}\]

@@

Filling the sub-goal with $\apply\: \V{D}\: \V{X}\: \V{Y}\:
\V{\sigma}$ completes the implementation. To sum up, we have
implemented the free monad construction for an arbitrary tagged
description. This gives the developer the ability, for any datatype,
to extend it with a notion of variable. Then, we have equipped this
structure with the corresponding monadic operations, \bind and
\return. This construction is an example of type-indexed
datatype~\cite{hinze:generic-haskell}, as found in Generic Haskell:
from a datatype, we build a new datatype and equip it with its
structure.

### 2010-06-30 papers/icfp-2010-desc/paper_discussion.tex
Desc@ICFP: Fix typos.

As such, our type theory suffers from an inconsistency. Indeed, the
typing rule $\Bhab{\Set}{\Set}$ leads to Girard's paradox. We made
that choice for presentational convenience, as universe stratification
is orthogonal to our work. Nonetheless, our universe of description
stratifies naturally. $\SYMBIDesc$ is self-encoding only in a
level-polymorphic sense.  Unsurprisingly, $\SYMBIDesc$ at level $l$ is
of type $\Set^{\blue{l+1}}$. Similarly, the interpretation of
$\SYMBIDesc$ at level $l$ is an object of type $\Set^{\blue{l}}$:
%
\[\stk{\begin{array}{@{}ll}
\SYMBIDesc^{\blue{l}} (\Bhab{\M{I}}{\Set^{\blue{l+1}}}) & :\Set^{\blue{l+1}} \\
\DVar{(\Bhab{\M{i}}{\M{I}})} &: \SYMBIDesc^{\blue{l}}{\M{I}} \\
\DConst{(\Bhab{\M{A}}{\Set^{\blue{l}}})} &:\SYMBIDesc^{\blue{l}}{\M{I}}       \\
\DProd{(\Bhab{\M{D}}{\SYMBIDesc^{\blue{l}}{\M{I}}})}{(\Bhab{\M{D}}{\SYMBIDesc^{\blue{l}}{\M{I}}})}
  & :\SYMBIDesc^{\blue{l}}{\M{I}}       \\
\DSigma{(\Bhab{\M{S}}{\Set^{\blue{l}}})}{(\Bhab{\M{D}}{\M{S}\To\SYMBIDesc^{\blue{l}}{\M{I}}})}
& : \SYMBIDesc^{\blue{l}}{\M{I}}  \\
\DPi{(\Bhab{\M{S}}{\Set^{\blue{l}}})}{(\Bhab{\M{D}}{\M{S}\To\SYMBIDesc^{\blue{l}}{\M{I}}})}
& : \SYMBIDesc^{\blue{l}}{\M{I}}  \\
\end{array}\smallskip \\
\idescop{\_\:}{}{}^{\blue{l}} : _{\PI{\V{I}}{\Set^{\blue{l+1}}}} \SYMBIDesc^{\blue{l}}\:\V{I} \To (\V{I} \To \Set^{\blue{l}}) \To \Set^{\blue{l}}    \\
\ldots
}
\]
Crucially, the types of data stored in an
\(\SYMBIDesc^{\blue{l}}{\M{I}}\) all live no higher---we may store an
\(I\) and a \(\Set^{\blue{l}}\) in a \(\Set^{\blue{l+1}}\).  The code
for \(\SYMBIDesc^{\blue{l}}{\M{I}}\) is an element of
\(\SYMBIDesc^{\blue{l+1}}{\Unit}\), so there is a spiral, not a cycle.
We have checked the construction using Agda's universe polymorphism,
coding $\SYMBIDesc$ in itself and have proving the isomorphism
between the host and the embedded universes.

@@

Our approach is follow the polytypic programming style, as initiated
by PolyP~\cite{jansson:polyp}. Indeed, we build generic functions by
induction on pattern functors. Unlike PolyP, we do not have to resort
to preprocessing: our datatypes are, natively, nothing but codes.

We share with Generic Haskell the \emph{type-indexed datatype}
approach~\cite{hinze:generic-haskell}, as exemplified by the free
monad construction: from datatype, we can compute new datatypes and
equip them with their structure. Generic Haskell also features
\emph{generic views}~\cite{holdermans:generic-view}, transparently
transforming the structure of datatype definitions. An example is the
tagged descriptions, presenting datatypes under a sum-of-sigmas
angle. Unlike Generic Haskell, we do not have to modify the compiler
to obtain views on datatypes: we can massage descriptions from inside
our language.

@@

Another generic programming paradigm is Scrap Your
Boilerplate~\cite{spj:syb} (SYB). Our proposal is different in various
ways. The corner stone of SYB is the \emph{spine} view of datatype
constructors. A piece of data is a spine composed by a constructor
applied to some arguments. SYB provides a combinator library to write
generic functions over spines. This relies on a $\CN{Typeable}$
type-class, allowing dynamic dispatch to datatype-specific
operations. As a result, SYB is not reflexive: it is restricted to
datatypes instanciating $\CN{Typeable}$. Moreover, it is limited to
building generic functions, hence type-indexed datatypes cannot be
implemented in this framework.

@@

However, we differ in several ways. First, there approach is
generative: each universe extends the base type theory with both type
formers and elimination rules. Thanks to levitation, we only rely on a
generic induction and a specialised $\SYMBswitchD$. Second,
the authors do not tackle the issue of \emph{programming} with
codes. We have shown how to abstract away codes and give a convenient
presentation to the developer. The authors often resort to an
extensional equality, while we have given an equality-agnostic
presentation. Beside, our universes are arranged so as to use
definitional equality as much as possible. Hence, in practice, the
developer is relieved from many proof obligations.

>>>>

As such, our type theory suffers from an inconsistency. Indeed, the
typing rule $\Bhab{\Set}{\Set}$ leads to Girard's paradox. We made
that choice for presentational convenience, as universe stratification
is orthogonal to our work. Nonetheless, our universe of description
stratifies naturally. $\SYMBIDesc$ is self-encoding only in a
level-polymorphic sense.  Unsurprisingly, $\SYMBIDesc$ at level $l$ is
of type $\Set^{\blue{l+1}}$. Similarly, the interpretation of
$\SYMBIDesc$ at level $l$ is an object of type $\Set^{\blue{l}}$:
%
\[\stk{\begin{array}{@{}ll}
\SYMBIDesc^{\blue{l}} (\Bhab{\M{I}}{\Set^{\blue{l+1}}}) & :\Set^{\blue{l+1}} \\
\DVar{(\Bhab{\M{i}}{\M{I}})} &: \SYMBIDesc^{\blue{l}}{\M{I}} \\
\DConst{(\Bhab{\M{A}}{\Set^{\blue{l}}})} &:\SYMBIDesc^{\blue{l}}{\M{I}}       \\
\DProd{(\Bhab{\M{D}}{\SYMBIDesc^{\blue{l}}{\M{I}}})}{(\Bhab{\M{D}}{\SYMBIDesc^{\blue{l}}{\M{I}}})}
  & :\SYMBIDesc^{\blue{l}}{\M{I}}       \\
\DSigma{(\Bhab{\M{S}}{\Set^{\blue{l}}})}{(\Bhab{\M{D}}{\M{S}\To\SYMBIDesc^{\blue{l}}{\M{I}}})}
& : \SYMBIDesc^{\blue{l}}{\M{I}}  \\
\DPi{(\Bhab{\M{S}}{\Set^{\blue{l}}})}{(\Bhab{\M{D}}{\M{S}\To\SYMBIDesc^{\blue{l}}{\M{I}}})}
& : \SYMBIDesc^{\blue{l}}{\M{I}}  \\
\end{array}\smallskip \\
\idescop{\_\:}{}{}^{\blue{l}} : _{\PI{\V{I}}{\Set^{\blue{l+1}}}} \SYMBIDesc^{\blue{l}}\:\V{I} \To (\V{I} \To \Set^{\blue{l}}) \To \Set^{\blue{l}}    \\
\ldots
}
\]
Crucially, the types of data stored in an
\(\SYMBIDesc^{\blue{l}}{\M{I}}\) all live no higher---we may store an
\(I\) and a \(\Set^{\blue{l}}\) in a \(\Set^{\blue{l+1}}\).  The code
for \(\SYMBIDesc^{\blue{l}}{\M{I}}\) is an element of
\(\SYMBIDesc^{\blue{l+1}}{\Unit}\), so there is a spiral, not a cycle.
We have checked the construction using Agda's universe polymorphism,
coding $\SYMBIDesc$ in itself and have proved the isomorphism
between the host and the embedded universes.

@@

Our approach follows the polytypic programming style, as initiated
by PolyP~\cite{jansson:polyp}. Indeed, we build generic functions by
induction on pattern functors. Unlike PolyP, we do not have to resort
to preprocessing: our datatypes are, natively, nothing but codes.

We share with Generic Haskell the \emph{type-indexed datatype}
approach~\cite{hinze:generic-haskell}, as exemplified by the free
monad construction: from a datatype, we can compute new datatypes and
equip them with their structure. Generic Haskell also features
\emph{generic views}~\cite{holdermans:generic-view}, transparently
transforming the structure of datatype definitions. An example is the
tagged descriptions, presenting datatypes under a sum-of-sigmas
angle. Unlike Generic Haskell, we do not have to modify the compiler
to obtain views on datatypes: we can massage descriptions from inside
our language.

@@

Another generic programming paradigm is Scrap Your
Boilerplate~\cite{spj:syb} (SYB). Our proposal is different in various
ways. The corner stone of SYB is the \emph{spine} view of datatype
constructors. A piece of data is a spine composed by a constructor
applied to some arguments. SYB provides a combinator library to write
generic functions over spines. This relies on a $\CN{Typeable}$
type-class, allowing dynamic dispatch to datatype-specific
operations. As a result, SYB is not reflexive: it is restricted to
datatypes instantiating $\CN{Typeable}$. Moreover, it is limited to
building generic functions, hence type-indexed datatypes cannot be
implemented in this framework.

@@

However, we differ in several ways. First, their approach is
generative: each universe extends the base type theory with both type
formers and elimination rules. Thanks to levitation, we only rely on a
generic induction and a specialised $\SYMBswitchD$. Second,
the authors do not tackle the issue of \emph{programming} with
codes. We have shown how to abstract away codes and give a convenient
presentation to the developer. The authors often resort to an
extensional equality, while we have given an equality-agnostic
presentation. Beside, our universes are arranged so as to use
definitional equality as much as possible. Hence, in practice, the
developer is relieved from many proof obligations.
